\documentclass[10pt]{article}
\usepackage[top=0.6in,bottom=0.6in,left=0.72in,right=0.72in]{geometry}
\usepackage{booktabs}
\usepackage{amssymb,amsmath}
\usepackage[hidelinks]{hyperref}
\usepackage{parskip}
\usepackage{microtype}
\usepackage{titlesec}
\titlespacing*{\section}{0pt}{5pt}{1pt}
\titlespacing*{\subsection}{0pt}{4pt}{1pt}
\setlength{\parskip}{2pt}
\pagestyle{empty}

\begin{document}

\begin{center}
{\large\bfseries Reproducing Multi-Agent System Failures:
ChatDev~1.0 and Magentic-One under gpt-5.4-mini}\\[2pt]
MAS memory research --- 2-page summary --- June 10, 2026
\end{center}

\textbf{Question.} MAST (``Why Do Multi-Agent LLM Systems Fail?'',
arXiv:2503.13657) groups multi-agent failures into three categories;
category~2, \emph{inter-agent misalignment}, covers failures of communication
between agents --- information withheld, input ignored, conversations reset.
We ask whether these are \emph{capability} failures (they fade with a
stronger model) or \emph{structural} failures (built into how the systems
communicate and remember). This report covers step one: re-run a sample of
MAST's tasks with a two-generation-newer model and re-label the runs with a
better judge.

\section{Dataset construction}
\textbf{The 30-task dataset was deliberately selected to maximize the chance
of observing inter-agent misalignment; it is not a representative sample, and
absolute failure rates should not be read as benchmark scores.} MAST's public
per-trace labels turned out to be unusable (a join error assigns labels by
row position), so we rebuilt ground truth: ChatDev human annotations
recovered from the MAST repo's git history, and Magentic-One outcomes
computed against the GAIA gold answers. We then read each candidate's
\emph{original GPT-4o trace} and kept it only if the original failure
involved agent-to-agent communication (rather than one agent simply getting
the answer wrong). Final set: 15 ChatDev game-building tasks
and 15 Magentic-One GAIA questions, each mixing high/medium/low-likelihood
cases plus originally-\emph{solved} controls (3 and 2 respectively).

\section{The two architectures}
\textbf{ChatDev 1.0} simulates a software company: a fixed ``waterfall'' of
phases (requirements $\rightarrow$ coding $\rightarrow$ three review rounds
$\rightarrow$ testing $\rightarrow$ docs), each phase a dialogue between two
role-prompted copies of the model (CEO, Programmer, Reviewer, Tester).
What passes from one phase to the next is \emph{only the project's code
files} --- and those files are created indirectly: when the Programmer
``writes code'', it types it into a chat message as filename-tagged code
blocks, and after the message a parser extracts each block and saves it to
the named file. Two details are load-bearing: (i)~each reply has a hard
length cap ($4096$ tokens minus the prompt), so a long file gets cut off
mid-line, and nothing checks for that; (ii)~the parser only saves properly
closed code blocks --- a cut-off block is \emph{silently skipped}. So an
agent can ``write a file'', see no error, and the file never comes to
exist. \textbf{Magentic-One} is a hub-and-spoke system for web research:
an Orchestrator holds a ``ledger'' (collected facts + current plan),
delegates one step at a time to four specialists (a screenshot-based
WebSurfer, FileSurfer, Coder, ComputerTerminal), and when progress stalls it
rewrites the ledger from its own summary of history --- a built-in reset.

\emph{Reproduction setup.} Both systems run their official, unmodified code
(ChatDev v1.1.6; Magentic-One via the MAST authors' own harness, AutoGen
0.4.8). The single change is the model behind the API: a local proxy
transparently serves \texttt{gpt-5.4-mini} where the systems believe they
are calling GPT-4o (tool calls, JSON mode, and screenshot vision verified
working). Deviations: code runs on the host rather than in Docker, and some
GAIA answers have drifted since 2024 because the live web changed.

\section{Outcomes vs.\ the original GPT-4o runs}
Magentic-One: $2/15$ solved originally $\rightarrow$ $3/15$ now (one failure
flipped to pass, no regressions, both controls still pass). ChatDev: $3/15
\rightarrow 5/15$ (two flips; all three controls still pass; caveat: original
grades are human annotations, ours come from the judge). \textbf{The
selected failure set largely survives the model upgrade}, which is what the
structural hypothesis predicts.

\section{MAST taxonomy statistics}
Every trace is labeled by a two-tier gpt-5.5 judge: tier~1 reads the full
trace with \emph{no taxonomy} and reports anything problematic in free form;
tier~2 then assigns the 14 MAST modes, and every flag must carry a verbatim
quote from the trace or it is dropped. Calibrating against the
human-annotated traces: modes where an agent visibly does the wrong thing
(2.3--2.6, 3.x) are detected reliably --- including 2.4 and 2.5, which
MAST's own o1-based judge had emitted \emph{zero} times --- while modes
defined by an \emph{absence} of action (2.1, 2.2) are caught only about half
the time, so those two rows below are lower bounds.

\begin{center}\small\renewcommand{\arraystretch}{0.95}%
\begin{tabular}{llcc}\toprule
Mode & Name & ChatDev (n=15) & Magentic-One (n=15)\\\midrule
2.1 & Conversation Reset & 7 & 4\\
2.2 & Fail to Ask for Clarification & 5 & 8\\
2.3 & Task Derailment & 6 & 8\\
2.4 & Information Withholding & \textbf{15} & 4\\
2.5 & Ignored Other Agent's Input & \textbf{15} & \textbf{13}\\
2.6 & Action--Reasoning Mismatch & 12 & 13\\
3.1--3.3 & Verification failures & 13/13/13 & 13/13/13\\
\bottomrule
\end{tabular}
\end{center}

The two systems show \emph{opposite} category-2 profiles, each matching its
architecture. ChatDev flags Information Withholding in all 15 runs ---
including the three that produced working software --- because its
code-persistence channel loses files in every run; whether the task succeeds
depends only on \emph{which} files are lost. Magentic-One shows little
withholding (the hub sees every message, so information rarely fails to
arrive) but pervasive ignored input (the hub or a specialist receives
information and doesn't act on it). That incidence tracks topology rather
than model is the core evidence for the structural reading.

\section{Open-ended judge trends}
The taxonomy-free tier produced $\sim$650 findings; three mechanism chains
dominate. (1)~\emph{ChatDev's lossy persistence chain}: code truncated
$\rightarrow$ file silently dropped $\rightarrow$ reviewer repeats the
identical complaint $\rightarrow$ tester approves anyway (the loss step
appears in the first case below; full chain in the DouDizhuPoker brief of
\texttt{main.pdf}). (2)~\emph{ChatDev error masking}: when a missing
file finally crashes the build, the Programmer patches \emph{around} the
crash (try/except fallbacks) instead of regenerating the file it already
wrote once. (3)~\emph{Magentic-One unsupported finals}: the Orchestrator
answers from memory or partial evidence instead of completing its own plan
(7 traces), and ledger rewrites discard progress already made.

\section{Four case briefs (chosen for inter-agent misalignment relevance)}
\small

\subsection*{ChatDev: Gomoku --- \emph{succeeded}, yet the communication channel failed}
Gomoku (five-in-a-row) is a \emph{control}: the original GPT-4o run also
solved it, and ours does too. Even so, the file-saving channel of \S2 failed
mid-run: the Programmer produced three correct files, but the parser
attached the wrong block to one of them, saving the board-logic file
\texttt{board.py} with the literal one-line content ``\texttt{python
main.py}'' (a shell command from the same message). A correct implementation
was produced and destroyed in transit --- Information Withholding (2.4)
without any agent deciding to withhold anything. Recovery succeeded here:
the Reviewer flagged the garbage file, and the regenerated \texttt{board.py}
was short enough to survive, so the finished game works. This is why 2.4
appears in all 15 ChatDev runs including the successful ones: the loss
mechanism fires regardless, and the task outcome depends only on whether the
lost piece happens to be recoverable in the remaining rounds.

\subsection*{ChatDev: Wordle --- failed; behavioral misalignment, no truncation involved}
The task: a terminal Wordle clone (guess a daily five-letter word in six
tries, with colored feedback). In contrast to the case above, nothing here
is lost in transit --- each failure is an agent's own decision, and the run
still fails. (1)~The prompt explicitly demands a terminal game, but
ChatDev's pipeline injects a boilerplate instruction that the software
``should be equipped with graphical user interface''; facing two
contradictory requirements, the Programmer silently picked one and moved on
rather than asking which was meant (2.2 Fail to Ask for Clarification ---
the same mode the MAST human annotators marked on the original GPT-4o run
of this task). (2)~The Reviewer gave a concrete, correct instruction: the
word list contains the six-letter word ``debate'', remove it. The
Programmer went through the modification round, yet the shipped code still
contains ``debate'' --- feedback received and not acted on (2.5), leaving
the delivered game in violation of the five-letter spec. (3)~Another review
round declared the duplicate-letter scoring the highest-priority bug while
recommending, as the fix, the exact code already present (2.6
Action--Reasoning Mismatch: the reviewer's diagnosis and proposal
contradict each other), sending the Programmer to churn on a non-bug.
(4)~The automated test ran the interactive game with no keyboard attached
and crashed on \texttt{EOFError}; the ``fix'' made the program exit
immediately whenever input is unavailable, after which the test passed
\emph{because gameplay was never reached} (3.3): the system made the test
pass rather than the game work.

\subsection*{Magentic-One 5a0c1adf --- failed; correct plan, specialist doesn't execute}
A GAIA web-research question: among winners of the Malko conducting
competition after 1977, find the one whose recorded nationality is a country
that no longer exists. The Orchestrator planned correctly and gave WebSurfer
specific, correct instructions --- open the winners archive on the
competition's official site. WebSurfer instead kept re-running general web
searches that hit CAPTCHAs and dead ends, and never opened the archive page
it was pointed to: the instruction arrived and was ignored (2.5). After the
loop stalled, the Orchestrator broke its own stated rule (``do not guess
before filtering all winners'') and submitted an unsupported guess (2.6).
The information flow worked; execution and verification did not.

\subsection*{Magentic-One 72e110e7 --- failed; reset loop replaces recovery}
A bibliographic lookup in BASE, an academic search engine. The needed page
rendered blank in the browser --- a recoverable tool failure. WebSurfer
responded only by waiting, despite explicit instructions to inspect the page
source or try another route (2.5). The Orchestrator then fell into
Magentic-One's built-in reset behavior: it kept rewriting its fact sheet and
plan from scratch (2.1 Conversation Reset) --- and although the rewritten
plans correctly named the fix (use an archived copy of the page), it never
delegated that step to any agent (2.4: the hub kept its own recovery plan to
itself) before answering ``unknown''. A minor tool hiccup became task
failure because the recovery mechanism resets instead of acting. \emph{(All
30 case briefs: \texttt{main.pdf}.)}

\end{document}
